To Airbnb?: Factors impacting short-term leasing preference

### 2018-11-09 papers/full_paper.Rmd
Revision, Round 2, Iteration 2

---
title: "To Airbnb?: A look at Long vs Shorm Term Leasing"
subtitle: "Andy Krause and Gideon Aschwanden"
author: "Zillow and University of Melbourne"
date: "Revised: November 10th, 2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
editor_options: 
  chunk_output_type: console
---

@@

@@

The growth of Airbnb and other short-term rental platforms have presented absentee owners of urban residential properties with a choice of leasing strategy: traditional long-term rental or a short-term approach ("Airbnb-ing").  In this paper we identify those situations -- location, structure type and property characteristics -- that lead to the highest likelihood of favoring a short-term strategy over a long term one. We also find that hosting policies matter, suggesting that even the right property may need the right owner(s) or strategy to make short-term rental the more profitable approach.

@@

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating urban residential property. Should they though? Or rather, is it more profitable to do so when compared to traditional long-term rental (LTR)? Are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one? And, finally, do host policies matter?

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term urban residential rental markets in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. To avoid confounding market dynamics caused by traditional vacation rentals (beach, ski and other recreation) we limit our study area to the urbanized areas of Melbourne. Additionally, for any individual property the choice between short and long term leasing may be influenced by individual financial situations (tax and mortgage financing) as well as non-financial concerns such as time availability and personal attachement to the property.  While we analyze individual properties, we acknowledge that there are many unobserved considerations and, therefore, we generalize our findings to market segments, locations and host policy decisions.

@@

Our results show that the feasibility of a greater revenues through a short-term rental -- what we have termed ``Short-Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner or investor of an urban residential property should opt for a short-term leasing strategy is ''it depends''.

@@

Focusing specifically on the short-term rental market, @cassidy2007tourist suggest that within Australia's Gold Coast the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers (RUMs) and are often under-priced as there is little accountability to owners.  The authors also note that very little research has been put toward understanding this growing market segment.

A collection of recent work has looked into the impact of short-term markets on housing prices and affordability. These working papers examine the influence of Airbnb on existing housing prices in New York [@sheppard2016airbnb], Los Angeles [@lee2016airbnb] and Boston [@merante2016home]. Though very broad analyses, all suggest that Airbnb has had a positive impact on local home prices and/or a negative impact on affordability. Other research examines the impact of short-term rentals on the hotel market, with @gutierrez2016airbnb finding that Airbnb and hotels have similar location patterns in Barcelona and @zervas2014rise showing an inverse relationship between Airbnb listings and hotel revenues.

@@

Though issues surrounding second homes have been primarily treated as a rural phenomenon for decades  [@gallent2007second], a wave of ``new urban tourism'' [@fuller2014stop] has focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many urban residential properties are not permitted for or regulated as commercial entities, STRs can have serious effects on immediate neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short-term rentals, particulary through Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

@@

We approach answering the question of whether it is more profitable to pursue a short- or long-term leasing strategy by creating and assessing a scenario likely to be faced by a recent purchaser of an urban residential investment property.  More specifically, we examine the following case:

* An investor purchases a property in the urbanized area of Melbourne, Australia on September 1, 2015.
* They are looking to maximize the net revenue over the 12-month period from September 1, 2015 to August 31, 2016.
* They do not intend to inhabit the home themselves. 
* The choices are 1) traditional year-long rental (long-term), or: 2) short-term, nightly Airbnb-type rental.
* Only direct monetary revenues and costs are considered, others such as liquidity and inconvenience are ignored as these concerns are influenced by owner situations, not property characteristics or hosting policies.  
* As most long-term rentals in Melbourne use a leasing agency, we assume use of an agency.

Australia implements a particular tax policy, often termed `negative gearing', that allows for an income tax deduction to real estate investors who experience greater costs than revenues on investment properties.^[See @fane2004negative for a complete explanation of negative gearing.] To date, this policy only applies to long-term rental properties.  While this policy can influence the net revenue (post tax) of a property, the negative gearing policy requires knowledge of the investor's personal income situation.  As we do not have access to this information in the marketplace, we will not be considering any effects of this policy in this analysis.

One benefit of a short-term rental may be the ability to adjust the nightly rate over the course of the year to reflect changing demand and market trends. As future market trends are unknown at the time of purchase and, within the span of 12 months, the decision of long- vs short-term leasing strategy is asymetrically irreversable (can go from short to long, but not long to short), we make our comparisons using a fixed nightly (short-term) rate to respresent the most likely state of knowledge for our hypothetical investor on September 1, 2015. Both long-term rents and short-term rates will be based on the market conditions as of this date.

We do not directly observe properties that are utilized in both a short- and long-term leasing strategy. Therefore, comparisons are done based on estimated or imputed revenues. Most specifically, we will estimate the long-term revenue for observations of properties currently listed on a short-term rental portal (Airbnb). Initially, we also attempted to estimate short-term rental revenue for long-term rental properties but found the process of imputing occupancy rates and nightly rate structures too inaccurate to provide valid comparisons.

In this scenario we compare the observed 12-month revenues from short-term rental properties with their hypothetical long-term rental revenues.  We define revenue here as the gross revenue minus the direct costs associated with each of the two tenure possibilities. Costs which are equivalent between the two strategies are ignored. For short-term properties the net revenue, $RevN_{ST}$, is expressed as:

@@

As our scenario compares a hypothetical and particular owner over the course of a single year the tax $T$ and $F$ costs are identical between the two leasing strategies and therefore cancel out. General maintenance $M$ expenses may differ between the two strategies, and an argument could be made for either to be higher. Long-term renters spend more time in the property over the course of the year and bring their own appliances^[White goods are usually not supplied in the Australian rental market.] and furniture which may damage the home. They are also more likely to affix items to the structure.  Conversely, short-term rentals are usually furnished and the occupants have less incentive not to be evicted as their damage is not often uncovered until after they have left.^[Note that Airbnb's rating policy may help protect hosts as bad guests can be flagged on the platform.]  Both arguments are meritous and we were unable to find any evidence to support maintenance costs being higher in short- or long-term rentals. As a result we have made the assumption that the maintenance costs will be identical, and they too cancel out.

Leasing costs $L$ are the fees paid to the rental agency to find and manage tenants over the life of the long-term lease.  In Melbourne, agency rates generally range from 5\% to 7\% per annum.  We have assumed leasing costs of 6\% of the total gross revenue of the long-term lease contract.

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. The Airbnb platform serves as the source of our short-term rentals data. On Airbnb, in addition to the nightly rate that an occupant pays to the owner, there is also a separate 'cleaning fee' that is attached, as is a booking fee payable to Airbnb.^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay.] Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item in the Airbnb data (paid separately and in addition to the nightly rate) they can also be ignored in the comparative analysis below.

@@

Throughout this analysis we will refer to any situations where the short-term net revenue exceeds the long-term net revenue as 'Short-Term Preferable' or a situation of 'Short-Term Preference' (STP).  For the purposes of this study we consider this a binary position leading to a binary decision of which leasing strategy to follow.

@@

Data for this study were gathered from two different sources, one for the short-term rental market and one for the long-term rental market.  Short-term rental market observations are from the Airbnb online portal. Airbnb is the largest short-term rental hub in the Melbourne region. As opposed to similar sites such as VRBO.com and Homeaway.com which focus on vacation homes in tourist dominated areas such as beach towns and ski resorts, Airbnb offers considerable coverage in urban areas and, therefore, directly competes to a greater degree with more traditional long-term rentals than the other sites.^[In fact, our choice of study area was done so to specifically minimize the number of potential traditional vacation homes in our sample.] The data on the Airbnb market in Melbourne have been purchased from www.airdna.co, a data provider specializing in Airbnb data collection and analysis. The short-term data include two types of observations, property-level and nightly-per-property. Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period.^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year during its time on the portal.  The nightly data indicate the advertised price, the occupancy status and a unique reservation identifier.

@@

Data on the long-term rental market was provided by Australia Property Monitors (APM), a Fairfax Group company. In raw form this dataset includes an observation for each rental property listing for each time a detail of the listing changed. For example, if a property was listed for rent for one month and the price was dropped once each week there would be four observations in the dataset. These four observations are grouped by an event-specific identification number. This number is specific to the listing event, not the property. In each case the last observation signifies the final price of the rental contract. The long-term lease data cover the period from January 2014 to December 2015.

Unfortunately, there are no common identifiers between the two datasets as the Airbnb data does not include addresses. Both sets of data do include latitude and longitude, however the accuracy of the Airbnb coordinates is suspect as these are often purposefully jittered to protect property anonymity and, as a result, in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on imprecise two-dimensional coordinates. As a result of this, we are unable to directly match observations from the two datasets.

@@

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an Airbnb-type, short-term rental approach to generating income. To determine the likely LTR nightly rate for any given property we use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

Both anecdotal evidence and Airbnb's own marketing campaigns suggest that there are at least two types of hosts in the market: 1) Those that lease out their properties all the time in the name of profit maximization (Profit Seekers); and 2) Those that lease out their property very infrequently, for example when they are away on vacation (Opportunistic Sharers).  For the purposes of this research we are focused on Profit Seekers; those owners who would be faced with deciding between a long- and short-term leasing strategy.

If these are the two dominant forms of hosts in the market, then a distribution of properties by the percent of time which they are blocked (not listed on the portal) should be highly bimodal with Profit Seekers in one group and Opportunistic Sharers in another. The left hand panel of Figure 1 shows that this breakdown holds somewhat, with peaks of hosts at each end, but that there are also a considerable number of hosts who block their properties some of the time (25\% to 75\%) .

Based on the data it appears there may also be a third potential host type, that of the host who utilizies multiple platforms to list their short-term property, of which Airbnb may just be one. A potential method to identify these hosts is to look at the number of individual periods of blocked days.  A host using multiple platforms will likely have many small periods of blocked days on Airbnb as they will block out those days in which reservations are made on other platforms. We test to see if this is a meaningful metric by plotting the total count of blocked periods versus the percentage of blocked days in the right panel of Figure 1. We do find many hosts with a large number of blocked periods, but, unfortunately, no true pattern emerges.

@@

The Airbnb data provides limited information on the structural characteristics. The only overlapping fields that we have between the two datasets are bedrooms and bathrooms. Additionally, as the markets and structural components of houses and apartments differ, we bifurcate our analysis between the two property types from this point forward.

We begin by removing all properties that are missing bedroom or bathroom information or have a likely data error in these fields (more than 14 bedrooms or 10 bathrooms). For both short- and long-term rentals we find that most apartments have fewer than four bedrooms and houses fewer than five.  For both property types, the vast majority of properties have three or fewer bathrooms. We filter the data accordingly. Additionally, note that short-term rentals consider studio apartments to be 0 bed units while the long-term data considers these as 1 bedroom. We convert the 0 bed short-term units to 1 bed units. A few short-term properties indicate no baths, which is likely an error so we remove these observations from the data. Also, here we notice that the short-term data gives bathrooms in halves, while the long-term data presents baths in whole numbers.  We round the baths in the short-term data up to whole integers.

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short-term rental) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two. From this analysis we see that there are 8 combinations that make up most all properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2, 3/3, 4/2 and 4/3.  The four bedrooms options are only present in the house category. We filter the data to these combinations.

@@

We begin by eliminating any observations in the outer suburbs.  Very few short-term rentals were located in these areas and those that were have a much higher likelihood of being vacation homes; a group of properties uses that doesn't match well to our research question. While the Inner and Middle suburb designation is a useful starting place, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events.

@@

![Suburb Categories](C:/Code/research/airbnbMelbourne/images/sub_loc_map.png)

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we filter our data to observations in suburbs that contain at least three STR and LTR observations.  The boundary of the suburbs meeting these criteria compose our study area (Figure 3). The location of the short- and long-term rentals, colored by submarket, are shown in Figure 4.  It should be noted that the beach areas in our study are densely developed and represently, generally speaking, higher end residential suburbs, not vacation-based communities (as one might find much farther down the coast).

![Long Term Observations](C:/Code/research/airbnbMelbourne/images/ltr_loc_map.png)

@@

Second, many of the short-term properties have a significant number of blocked days (up to 255), days in which the property was not offered on Airbnb.  Here too, there is a problem with comparing observed revenues from a property with a 20\% block rate with a hypothetical year long revenue stream, especially if the property was leased out on other portals during those blocked days.  And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should not necessarily apply to the hypothetical investor in our examined scenario.  In short, in either case the actual observed short-term revenue would be biased on the low end compared to what a revenue maximizing owner would be likely able to accomplish.

@@

1. Actual ($REV_a$)}: Observed bookings multiplied by median nightly rate

2. Potential ($REV_p$)}: Observed bookings extrapolated to a full year plus imputed bookings during observed blocked days.  Our attempt here is to represent the potential revenue if the property had been listed on Airbnb all year long.  The imputed bookings are imputed by comparing the property specific performance during the non-blocked days to overall market performance on those days and then applying that relative ratio to the overall market performance on the extrapolated and blocked days.

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Potential' revenue metric to be a better estimate of market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties (Figure 6). Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year.  Actual short-term revenues range from neglible to highs of greater than \$60,000, with a collection of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is shifted slightly to the right due to our time extrapolation and blocked days imputation.

While this simple analysis does not control for location and product differences, it does highlight the fact that some profit seeking Airbnb owners are likely making greater revenues than in long-term rentals, but that many are not. The high number of very low yield Airbnb properties suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

@@

We then average the hedonic model and matched pairs prediction for each short term property to create an estimate of its long-term rental rate. As a validation check, we find some variation between the predictions from the two models for specific properties, however, the overall correlation between the two estimates is 97.6%.

@@

Overall, 57.3%^[If we just look at actual revenue, ignoring blocked days and properties with less than a full 12 months of listing, the figure is 42.5%] of the 2,415 short-term urban rental properties in our sample were more profitable as a short-term rental than it would have been if rented long-term over the course of the year. While illustrative, a simple univariate analysis does not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below.

We specify three, increasingly complex, logistic regression models.

@@

  # XTable
  x.full <- xtable(full.table,
                  caption='Logistic Modeling Results (Likely Revenue)\n')
  align(x.full) <- 'lllll'

  # Print
  print(x.full,
       include.rownames=FALSE,
       hline.after=c(-1,0,15,20,20),
       caption.placement='top')

>>>>

---
title: "To Airbnb?: Factors impacting short-term leasing preference"
subtitle: "Andy Krause and Gideon Aschwanden"
author: "Zillow and University of Melbourne"
date: "Revised: November 10th, 2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
editor_options: 
  chunk_output_type: console
---

@@

  # library(knitr)
  # knit_hooks$set(plot = function(x, options) {
  # paste('<figure><figcaption>', options$fig.cap, '</figcaption><img src="',
  #       opts_knit$get('base.url'), paste(x, collapse = '.'),
  #       '"></figure>',
  #       sep = '')})

@@

The growth of Airbnb and other short-term rental platforms have presented absentee owners of urban residential properties with a choice of leasing strategy: traditional long-term rental or a short-term approach ("Airbnb-ing").  In this paper we identify those situations -- location, structure type and property characteristics -- that lead to the highest likelihood of favoring a short-term strategy over a long term one. Additionally, tests on the impacts of hosting policies suggest that even the right property may need the right owner(s) or strategy to make short-term rental the more profitable approach.

@@

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating urban residential property. Should they though? More specifically,  are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one? And, finally, do short-term hosting policies matter?

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term urban residential rental markets in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. To avoid confounding market dynamics caused by traditional vacation rentals (beach, ski and other recreation) we limit our study area to the urbanized areas of Melbourne. Additionally, for any individual property the choice between short and long term leasing may be influenced by individual financial situations (tax and mortgage financing) as well as non-financial concerns such as time availability and personal attachment to the property.  While we analyze individual properties, we acknowledge that there are many unobserved considerations and, therefore, we generalize our findings to market segments, locations and host policy decisions.

@@

Our results show that the feasibility of a greater revenues through a short-term rental -- what we have termed ''Short-Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner of an urban residential property is likely to earn more through a short-term leasing strategy than a long-term one is ''it depends''.

@@

Focusing specifically on the short-term rental market, @cassidy2007tourist suggest that within Australia's Gold Coast the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers and are often under-priced as there is little accountability to owners.  The authors also note that very little research has been put toward understanding this growing market segment.

A collection of recent work has looked into the impact of short-term markets on housing prices and affordability. This research examines the influence of Airbnb on existing housing prices nationally [@barron2018sharing] as well as focused work in New York [@sheppard2016airbnb], Los Angeles [@lee2016airbnb] and Boston [@merante2017home]. Though very broad analyses, all suggest that Airbnb has had a positive impact on local home prices and/or a negative impact on affordability. Other research examines the impact of short-term rentals on the hotel market, with @gutierrez2016airbnb finding that Airbnb and hotels have similar location patterns in Barcelona and @zervas2014rise showing an inverse relationship between Airbnb listings and hotel revenues.

@@

Though issues surrounding second homes have been primarily treated as a rural phenomenon for decades  [@gallent2007second], a wave of ``new urban tourism'' [@fuller2014stop] has focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many urban residential properties are not permitted for or regulated as commercial entities, STRs can have serious effects on immediate neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable; @gurran2018urban].  Many of the key legal decisions regarding short-term rentals, particulary through Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

@@

We approach answering the question of which property, location and hosting factors are related to Short Term Preference by gathering market evidence from both the short and long term rental markets. Using this market data, we compare the actual short term revenues of known Airbnb properties to their expected long-term rental revenues had they employed a long-term strategy.  Properties whose revenues were higher with a short-term strategy are considered to have shown Short-Term Preference (STP).  We then use this binary classification to construct statistical models aimed at assessing the impacts of property, location and hosting factors on the likelihood of Short-Term Preference.  Our study examines the Melbourne, Australia market over the September 1, 2015 to August 31, 2016 time frame.

The following assumptions are made:

1.  Only direct monetary revenues and costs are considered, others such as liquidity and inconvenience are ignored as these concerns are influenced by owner situations, not property characteristics or hosting policies.

2. As most long-term rentals in Melbourne use a leasing agency, we assume use of an agency.

3. Australia implements a particular tax policy, often termed `negative gearing', that allows for an income tax deduction to real estate investors who experience greater costs than revenues on investment properties.^[See @fane2004negative for a complete explanation of negative gearing.] To date, this policy only applies to long-term rental properties.  While this policy can influence the net revenue (post tax) of a property, the negative gearing policy requires knowledge of the investor's personal income situation.  As we do not have access to this information in the marketplace, any impacts from tax considerations are ignored.

One benefit of a short-term rental may be the ability to adjust the nightly rate over the course of the year to reflect changing demand and market trends. As future market trends are unknown at the time of purchase and, within the span of 12 months (standard lease length), the decision of long- vs short-term leasing strategy is asymetrically irreversable (can go from short to long, but not long to short), we make our comparisons using a long term rental rate this is fixed as of September 1, 2015 but a short-term nightly rate that varies with the market over the course of the year.

Unfortunately, we cannot not directly observe properties that are utilized in both a short- and long-term leasing strategy. Therefore, comparisons are done with actual short-term but estimated or imputed long-term revenues. Initially, we also attempted to estimate short-term rental revenue for long-term rental properties but found the process of imputing occupancy rates and nightly rate structures too inaccurate to provide valid comparisons.

For our comparisons we define (net) revenue as the gross revenue minus the direct costs associated with each of the two tenure possibilities. Costs which are equivalent between the two strategies are ignored. For short-term properties the net revenue, $RevN_{ST}$, is expressed as:

@@

As our scenario compares a hypothetical and particular owner over the course of a single year the tax, $T$, and, $F$, costs are identical between the two leasing strategies and therefore cancel out. General maintenance, $M$, expenses may differ between the two strategies, and an argument could be made for either to be higher. Long-term renters spend more time in the property over the course of the year and bring their own appliances^[White goods are usually not supplied in the Australian rental market.] and furniture which may damage the home. They are also more likely to affix items to the structure.  Conversely, short-term rentals are usually furnished and the occupants have less incentive not to be evicted as their damage is often not uncovered until after they have left.^[Note that Airbnb's rating policy may help protect hosts as bad guests can be flagged on the platform.]  Both arguments are meritous and we were unable to find any evidence to support maintenance costs being higher in short- or long-term rentals. As a result we have made the assumption that the maintenance costs will be identical, and they too cancel out.

Leasing costs, $L$, are the fees paid to the rental agency to find and manage tenants over the life of the long-term lease.  In Melbourne, agency rates generally range from 5\% to 7\% per annum.  We have assumed leasing costs of 6\% of the total gross revenue of the long-term lease contract.

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. The Airbnb platform serves as the source of our short-term rentals data. On Airbnb, in addition to the nightly rate that an occupant pays to the owner, there is also a separate 'cleaning fee' that is attached, as is a booking fee payable to Airbnb.^[This fee could be considered analogous to the leasing fees, $L$, that long-term rentals pay.] Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item in the Airbnb data (paid separately and in addition to the nightly rate) they can also be ignored in the comparative analysis below.

@@

Throughout this analysis we will refer to any situations where the short-term net revenue exceeds the long-term net revenue as 'Short-Term Preferable' or a situation of 'Short-Term Preference' (STP).  For the purposes of this study we consider this a binary position.

@@

Data for this study were gathered from two different sources, one for the short-term rental market and one for the long-term rental market.  Short-term rental market observations are from the Airbnb online portal. Airbnb is the largest short-term rental hub in the Melbourne region. As opposed to similar sites such as VRBO.com and Homeaway.com which focus on vacation homes in tourist dominated areas such as beach towns and ski resorts, Airbnb offers considerable coverage in urban areas and, therefore, directly competes to a greater degree with more traditional long-term rentals than the other sites.^[In fact, our choice of study area was done so to specifically minimize the number of potential traditional vacation homes in our sample.] The data on the Airbnb market in Melbourne have been purchased from www.airdna.co, a data provider specializing in Airbnb data collection and analysis. The data offers both nightly observations indicating the advertised price, the occupancy status and a unique reservation identifier as well as general property and host policy information.

@@

Data on the long-term rental market was provided by Australia Property Monitors (APM), a Fairfax Group company. The long-term lease contains both property specific information (beds, baths, location, etc.) as well as a history of listing prices.  These data cover the period from January 2014 to December 2015.

Unfortunately, there are no common identifiers between the two datasets as the Airbnb data does not include addresses. Both sets of data do include latitude and longitude, however the accuracy of the Airbnb coordinates is suspect as these are often purposefully jittered to protect property anonymity and, as a result, in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on imprecise two-dimensional coordinates. As a result, we are unable to directly match observations from the two datasets.

@@

We begin by filtering observations by time. Our study design treats September 1, 2015 as the 'decision date' on which a property owner would decide between a long or short term strategy.  To determine the likely LTR nightly rate for the subsequent 12 months we use the observed rental transactions from September 1, 2014 to August 31, 2015 to build a model to estimate market rent as of August 31, 2015. A short term leasing approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this example we will use the observed Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

Both anecdotal evidence and Airbnb's own marketing campaigns suggest that there are at least two types of hosts in the market: 1) Those that lease out their properties all the time in the name of profit maximization (Profit Seekers); and 2) Those that lease out their property very infrequently, for example when they are away on vacation (Opportunistic Sharers).  For the purposes of this research we are focused on the Profit Seekers; the properties mostly likely to be available for long or short term rentals.  
If these are the two dominant forms of hosts in the market, then a distribution of properties by the percent of time which they are blocked (not listed on the portal) should be highly bimodal with Profit Seekers in one group and Opportunistic Sharers in another. The left hand panel of Figure 1 shows that this breakdown holds somewhat, with peaks of hosts at each end, but that there are also a considerable number of hosts who block their properties some of the time (25\% to 75\%) .

Based on the data it appears there may also be a third potential host type, that of the host who utilizies multiple platforms to list their short-term property, of which Airbnb may just be one. A potential method to identify these hosts is to look at the number of individual periods of blocked days.  A host using multiple platforms will likely have many small periods of blocked days on Airbnb as they will block out those days in which reservations are made on other platforms. We test to see if this is a meaningful metric by plotting the total count of blocked periods versus the percentage of blocked days in the right panel of Figure 1. We do find many hosts with a large number of blocked periods, but, also many with few and, unfortunately, no true pattern emerges.

@@

The Airbnb data provides limited information on the structural characteristics. The only overlapping fields that we have between the two datasets are structure type, bedrooms and bathrooms. Additionally, as the markets and structural components of houses and apartments differ, we bifurcate our analysis between the two property types from this point forward.

We begin by removing all properties that are missing bedroom or bathroom information or have a likely data error in these fields (more than 14 bedrooms or 10 bathrooms). For both short- and long-term rentals we find that most apartments have fewer than four bedrooms and houses fewer than five.  For both property types, the vast majority of properties have three or fewer bathrooms. We filter the data accordingly.^[Additionally, note that short-term rentals consider studio apartments to be 0 bed units while the long-term data considers these as 1 bedroom. We convert the 0 bed short-term units to 1 bed units.] A few short-term properties indicate no baths, which is likely an error so we remove these observations from the data. Also, here we notice that the short-term data gives bathrooms in halves, while the long-term data presents baths in whole numbers.  We round the baths in the short-term data up to whole integers.

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short-term rental) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two. From this analysis we see that there are eight combinations that make up the bulk of the properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2, 3/3, 4/2 and 4/3.  The four bedrooms options are only present in the house category. We filter the data to these combinations.

@@

We begin by eliminating any observations in the outer suburbs.  Very few short-term rentals were located in these areas and those that were have a much higher likelihood of being vacation homes; a group of properties uses that doesn't match well to our research question. While the Inner and Middle suburb designation is a useful starting place, we belief that a few additional spatial features can also influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events.

@@

<!-- ![Suburb Categories](C:/Code/research/airbnbMelbourne/images/sub_loc_map.png) -->

We assign submarket designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-3 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we filter our data to observations in suburbs that contain at least three STR and LTR observations.  The boundary of the suburbs meeting these criteria compose our study area. The location of the short- and long-term rentals, colored by submarket, are shown in Figure 4.  It should be noted that the beach areas in our study are densely developed and represently, generally speaking, higher end residential suburbs, not vacation-based communities as one might find much farther south along the coast.

![Study Area and Observation Locations](C:/Code/research/airbnbMelbourne/images/ltr_loc_map.png)

@@

Second, many of the short-term properties have a significant number of blocked days (up to 255), days in which the property was not offered on Airbnb.  Here too, there is a problem with comparing observed revenues from a property with a 20\% block rate with a hypothetical year long revenue stream, especially if the property was leased out on other portals during those blocked days.  And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should factor into the comparative analyses we make.  In short, in either case the actual observed short-term revenue would be biased on the low end compared to what a revenue maximizing owner would be likely able to accomplish.

@@

1. Actual ($REV_a$): Observed bookings multiplied by median nightly rate

2. Potential ($REV_p$): Observed bookings extrapolated to a full year plus imputed bookings during observed blocked days.  Our attempt is to represent the potential revenue if the property had been listed on Airbnb all year long.  The imputed bookings are calculated by comparing the property specific performance during the non-blocked days to overall market performance on those days and then applying that relative ratio to the overall market performance on the extrapolated and blocked days.

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Potential' revenue metric to be a better estimate of market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties (Figure 6). Revenues from the long-term rentals are far more consistent, with the vast majority totaling between A\$12,000 to A\$30,000 per year.  Actual short-term revenues range from neglible to highs of greater than A\$60,000, with a collection of the properties showing very limited (less than A\$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is shifted slightly to the right due to our time extrapolation and blocked days imputation. While this simple analysis does not control for location and product differences, it does highlight the fact that some profit seeking Airbnb owners are likely making greater revenues than in long-term rentals, but that many are not.

@@

We then average the hedonic model and matched pairs prediction for each short term property to create an estimate of its long-term rental rate. As a validation check, we find some variation between the predictions from the two models for specific properties, however, the overall correlation between the two estimates is 89.4%.

@@

Overall, 57.3%^[If we just look at actual revenue, ignoring blocked days and properties with less than a full 12 months of listing, the figure is 42.5%] of the 2,415 short-term urban rental properties in our sample were more profitable as a short-term rental than it would have been if rented long-term over the course of the year. While illustrative, a simple univariate analysis does not fully control for location, property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of three increasingly complex, logistic regression models.

@@

  xf_cap <- paste0('Dependent Variable = Short Term Preference.  These three logistic',
                   ' regression models use 2,415 short term rental observations to ',
                   'examine the impacts of various features on short term preference (',
                   ' the likelihood of greater net revenues from a short term vs a long',
                   ' term leasing strategy).')
